---
title: "1.extract_fluxes"
author: "abbey yatsko"
date: "26/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Extract fluxes for termite mounds from LGR raw data 

```{r workspace prep}
library(data.table)
library(scales)
library(ggplot2)
library(ggpubr)
library(sqldf)
library(dplyr)
library(tidyverse)
library(stringr)
library(plyr)
```

## 1. May 2022 data load and prep
* may 2022 has wonky formatting due to LGR time clock error 

```{r data prep}
# all LGR files loading in
dataTMI <- list.files(path ="/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/may22LGR", pattern='f00', full.names= T) %>%
  setNames(nm = .) %>% 
  lapply(read.csv,skip=1) %>%
  bind_rows(.id = "file") %>%
  mutate(file = str_extract(file, "f00\\d+"))

# there is a pgp error message at the end of some LGR files. need to remove these lines 
# remove any entry where there is NA for both x.CH4._ppm and x.CH4._ppm_sd
data_clean <- subset(dataTMI, X.CH4._ppm!="NA")

# read in metadata to alter dates for each f#### file so that the fDOY makes sense 
# internal time clock in the LGR was broken at the time of use, so dates need to be manipulated
meta_TMI <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/TMI_flux_metadata_may22.csv")
file <- unique(meta_TMI$file)
file_years <- as.data.frame(file)

# create repeating months 1-12 to make fDOY numbers have a smaller range 
file_years$month_alt <- rep(01:12, length.out=67)

# create new years so that each date is unique and not repeated for the fDOY
file_years$year_alt <- ""
file_years$year_alt[1:12] <- 2002
file_years$year_alt[13:24] <- 2003
file_years$year_alt[25:36] <- 2004
file_years$year_alt[37:48] <- 2005
file_years$year_alt[49:60] <- 2006
file_years$year_alt[61:67] <- 2007

# merge in alternative year and month to flux df
data_clean <- merge(data_clean, file_years, by = c("file")) 

# re-order month_alt year_alt 
data_clean <- data_clean %>% relocate(year_alt, .before = X.CH4._ppm)
data_clean <- data_clean %>% relocate(month_alt, .before = year_alt)
```

### Create alternative times 
* sub in alternative year and month to replace the broken time clock date/time
```{r alternative time}
# replace original year (always 2002) in date_time with the alternative year (needed for fDOY)
# remove 2002 year from Time column so that unique year can be pasted in
data_clean$Time <- gsub("2002", "", data_clean$Time)

# create altered time column by pasting in alternative year to Time column in the place of the old 2002
data_clean$time_altyr <- with(data_clean, paste(substr(Time, 1, 8),
                           year_alt,
                           substr(Time, 9, nchar(Time)),
                           sep = ""))

# remove first two characters from time_altmo column strings (for replacing month)
data_clean$time_altmo <- substring(data_clean$time_altyr , 5)

# paste in month_alt 
data_clean$time_alt <- with(data_clean, paste(substr(time_altmo, 1, 0),
                                   month_alt,
                                   substr(time_altmo, 1, nchar(time_altmo)),
                                   sep = ""))

# move up time_alt to compare with original Time
data_clean <- data_clean %>% relocate(time_alt, .before = month_alt)
```

### Format dates and times 
```{r datetime format}
# pull out date and time data - use time_alt
date_time <- strptime(data_clean[,3],format='%m/%d/%Y %H:%M:%S')

# formatting dates and times ----
# add year,month,day,JD,hour,min,sec columns to dataframe
Year <- as.numeric(format(date_time,'%Y'))
Month <- as.numeric(format(date_time,'%m'))
Day <- as.numeric(format(date_time,'%d'))
fDOY <- as.numeric(julian(date_time,'2002-01-01'))  #Change for year
Hour <- as.numeric(format(date_time,'%k'))
Min <- as.numeric(format(date_time,'%M'))
Sec <- as.numeric(format(date_time,'%S'))
data_clean <- cbind(date_time,Year,Month,Day,fDOY,Hour,Min,Sec,data_clean[,-3])
# save LGR data as data.table
data_clean <- data.table(data_clean)
```

### Preliminary visualization
```{r}
# for co2
co2 <- ggplot(data_clean, aes(date_time, X.CO2.d_ppm)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("CO2 flux")

# for ch4 
ch4 <- ggplot(data_clean, aes(date_time, X.CH4.d_ppm)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("CH4 flux")

# looks like there are the right number of lines that represent the 67 flux files 
```

### Merge with metadata
```{r}
# merge in year_alt and month_alt to metadata
meta_TMI <- merge(meta_TMI, file_years, by = c("file")) 

# add in column for measurement_day_alt
meta_TMI$measurement_day_alt <- "/01/"

# paste in year_alt to the end of measurement_day_alt
meta_TMI$measurement_day_altyr <- with(meta_TMI, paste(substr(measurement_day_alt, 1, 8),
                                              year_alt,
                                              substr(measurement_day_alt, 9, nchar(measurement_day_alt)),
                                              sep = ""))
# paste in month_alt to beginning of  measurement_day_alt
meta_TMI$measurement_day_alt <- with(meta_TMI, paste(substr(measurement_day_altyr, 1, 0),
                                              month_alt,
                                              substr(measurement_day_altyr, 1, nchar(measurement_day_altyr)),
                                              sep = ""))

# reposition measurement_day_alt in df 
meta_TMI <- meta_TMI %>% relocate(measurement_day_alt, .before = measurement_day_actual)
meta_TMI <- meta_TMI %>% relocate(year_alt, .before = measurement_day_actual)
meta_TMI <- meta_TMI %>% relocate(month_alt, .before = year_alt)

# format measurement_day_alt
as.Date(meta_TMI$measurement_day_alt, '%m/%d/%Y')
```

### Add start and end time for flux measurements 
```{r}
## Pull out date and time data
date_time_start <- strptime(paste(meta_TMI$measurement_day_alt,meta_TMI$flux_start),'%m/%d/%Y %H:%M:%S')

## Add 30 seconds to set buffered start time
date_time_start <- date_time_start+30

## Add 1.5 min (90 sec) to determine end time
date_time_end <- date_time_start+90

## Add fDOY columns for start and endtime to dataframe
fDOY_start <- as.numeric(julian(date_time_start,'2002-01-01'))  #Change for year
fDOY_end <- as.numeric(julian(date_time_end,'2002-01-01'))  #Change for year
meta_TMI <- cbind(fDOY_start,fDOY_end,meta_TMI)

# calculate average respiration temperature - add column
meta_TMI$avg_respT <- rowMeans(meta_TMI[ , c(13:16)], na.rm = TRUE)

# change sample to factor level variables 
meta_TMI$sample <- as.factor(meta_TMI$sample) 

# rename co2 and ch4 columns because naming format including '.' messes up the sql code 
data_clean <- data_clean %>%
  dplyr::rename(CH4 = X.CH4._ppm,
         CO2 = X.CO2._ppm, 
         CH4_dry = X.CH4.d_ppm,
         CO2_dry = X.CO2.d_ppm)
```

## 2. Nov 2022 data load and prep
```{r}
# workplace setup 
setwd("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/nov22LGR")
filenames <- list.files(pattern='f000',full.names=T)

# read in LGR datafiles to list and go from .txt to .csv 
data_s <- lapply(filenames,read.csv,skip=1)
data_s <- rbindlist(data_s)
data_s <- subset(data_s, X.CH4._ppm!="NA")
as.character(data_s$Time)

date_time <- strptime(data_s$Time,format='%d/%m/%Y %H:%M:%S')

# formatting dates and times ----
# add year,month,day,JD,hour,min,sec columns to dataframe
Year <- as.numeric(format(date_time,'%Y'))
Month <- as.numeric(format(date_time,'%m'))
Day <- as.numeric(format(date_time,'%d'))
fDOY <- as.numeric(julian(date_time,'2022-01-01'))  #Change for year
Hour <- as.numeric(format(date_time,'%k'))
Min <- as.numeric(format(date_time,'%M'))
Sec <- as.numeric(format(date_time,'%S'))
data_clean2 <- cbind(date_time,Year,Month,Day,fDOY,Hour,Min,Sec,data_s[,-1])
# save LGR data as dt
data_clean2 <- data.table(data_clean2)
```

### Merge with metadata
```{r}
# read in file with metadata
meta_TMI2 <- read.csv('/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/TMI_flux_metadata_nov22.csv')

as.Date(meta_TMI2$measurement_day_actual, '%m/%d/%Y')

## Pull out date and time data
date_time_start <- strptime(paste(meta_TMI2$measurement_day_actual,meta_TMI2$flux_start),'%m/%d/%Y %H:%M:%S')

## Add 30 seconds to set buffered start time
date_time_start <- date_time_start+30

## Add 90 sec to determine end time
date_time_end <- date_time_start+90

## Add fDOY columns for start and endtime to dataframe
fDOY_start <- as.numeric(julian(date_time_start,'2022-01-01'))  #Change for year
fDOY_end <- as.numeric(julian(date_time_end,'2022-01-01'))  #Change for year
meta_TMI2 <- cbind(fDOY_start,fDOY_end,meta_TMI2)

# calculate average respiration temperature - add column
meta_TMI2$avg_respT <- rowMeans(meta_TMI2[ , c(9:11)], na.rm = TRUE)

# change sample to factor level variables 
meta_TMI2$sample <- as.factor(meta_TMI2$sample) 

# rename co2 and ch4 columns because naming format including '.' messes up the sql code 
data_clean2 <- data_clean2 %>%
  dplyr::rename(CH4 = X.CH4._ppm,
         CO2 = X.CO2._ppm,
         CH4_dry = X.CH4.d_ppm,
         CO2_dry = X.CO2.d_ppm)
```

## 3. Aug 2023 data load and prep
```{r}
# workplace setup 
setwd("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/aug23LGR")
filenames <- list.files(pattern='f00',full.names=T)

# read in LGR datafiles to list and go from .txt to .csv 
data_s <- lapply(filenames,read.csv,skip=1)
data_s <- rbindlist(data_s)
data_s <- subset(data_s, X.CH4._ppm!="NA")
as.character(data_s$Time)

date_time <- strptime(data_s$Time,format='%d/%m/%Y %H:%M:%S')

# formatting dates and times ----
# add year,month,day,JD,hour,min,sec columns to dataframe
Year <- as.numeric(format(date_time,'%Y'))
Month <- as.numeric(format(date_time,'%m'))
Day <- as.numeric(format(date_time,'%d'))
fDOY <- as.numeric(julian(date_time,'2023-01-01'))  #Change for year
Hour <- as.numeric(format(date_time,'%k'))
Min <- as.numeric(format(date_time,'%M'))
Sec <- as.numeric(format(date_time,'%S'))
data_clean3 <- cbind(date_time,Year,Month,Day,fDOY,Hour,Min,Sec,data_s[,-1])

# save LGR data as dt
data_clean3 <- data.table(data_clean3)
```

### Merge with metadata
```{r}
# read in file with metadata
meta_TMI3 <- read.csv('/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/TMI_flux_metadata_aug.csv')

as.Date(meta_TMI3$measurement_day_actual, '%m/%d/%Y')

## Pull out date and time data
date_time_start <- strptime(paste(meta_TMI3$measurement_day_actual,meta_TMI3$flux_start),'%m/%d/%Y %H:%M:%S')

## Add 30 seconds to set buffered start time
date_time_start <- date_time_start+30

## Add 90 sec to determine end time
date_time_end <- date_time_start+90

## Add fDOY columns for start and endtime to dataframe
fDOY_start <- as.numeric(julian(date_time_start,'2023-01-01'))  #Change for year
fDOY_end <- as.numeric(julian(date_time_end,'2023-01-01'))  #Change for year
meta_TMI3 <- cbind(fDOY_start,fDOY_end,meta_TMI3)

# calculate average respiration temperature - add column
meta_TMI3$avg_respT <- rowMeans(meta_TMI3[ , c(9:11)], na.rm = TRUE)

# change sample to factor level variables 
meta_TMI3$sample <- as.factor(meta_TMI3$sample) 

# clip out MD56 - impartial measurement 
meta_TMI3 <- meta_TMI3[!(meta_TMI3$sample=="MD56"),]

# rename co2 and ch4 columns because naming format including '.' messes up the sql code 
data_clean3 <- data_clean3 %>%
  dplyr::rename(CH4 = X.CH4._ppm,
         CO2 = X.CO2._ppm,
         CH4_dry = X.CH4.d_ppm,
         CO2_dry = X.CO2.d_ppm)
```

## 4. Merge may + nov + aug LGR files with metadata 
* goal is to clip out samples and remove measurements for when the chamber was empty 
```{r}
# may 2022
# complete merge for respiration data and metadata (may)
data_merge <- sqldf("select data_clean.fDOY, data_clean.CO2_dry, data_clean.CH4_dry, meta_TMI.sample, meta_TMI.avg_respT, meta_TMI.flux_source, meta_TMI.position, meta_TMI.tag, meta_TMI.directon,meta_TMI.chamber, meta_TMI.chamber_SA_cm2, meta_TMI.soil_moisture, meta_TMI.fDOY_start 
from data_clean LEFT JOIN meta_TMI ON (data_clean.fDOY BETWEEN meta_TMI.fDOY_start AND meta_TMI.fDOY_end)")

# filter out rows that  have no value for 'sample' - this means that no mound was being recorded
data_merge <- data_merge[complete.cases(data_merge[,c("sample")]),]

# check to make sure all of the mounds are there (67 total)
summary(unique(data_merge$sample))

# complete merge for respiration data and metadata (nov)
data_merge2 <- sqldf("select data_clean2.fDOY, data_clean2.CO2_dry, data_clean2.CH4_dry, meta_TMI2.sample, meta_TMI2.avg_respT, meta_TMI2.flux_source, meta_TMI2.position, meta_TMI2.tag, meta_TMI2.directon, meta_TMI2.chamber, meta_TMI2.chamber_SA_cm2, meta_TMI2.fDOY_start 
from data_clean2 LEFT JOIN meta_TMI2 ON (data_clean2.fDOY BETWEEN meta_TMI2.fDOY_start AND meta_TMI2.fDOY_end)")

# use 'complete.cases'
data_merge2 <- data_merge2[complete.cases(data_merge2), ]

# check to make sure all of the samples are there (24 total)
summary(unique(data_merge2$sample))

# complete merge for respiration data and metadata (aug)
data_merge3 <- sqldf("select data_clean3.fDOY, data_clean3.CO2_dry, data_clean3.CH4_dry, meta_TMI3.sample, meta_TMI3.avg_respT, meta_TMI3.flux_source, meta_TMI3.position, meta_TMI3.tag, meta_TMI3.directon, meta_TMI3.chamber, meta_TMI3.chamber_SA_cm2, meta_TMI3.fDOY_start 
from data_clean3 LEFT JOIN meta_TMI3 ON (data_clean3.fDOY BETWEEN meta_TMI3.fDOY_start AND meta_TMI3.fDOY_end)")

# use 'complete.cases'
data_merge3 <- data_merge3[complete.cases(data_merge3), ]

# check to make sure all of the samples are there (25 total)
summary(unique(data_merge3$sample))
```

### Visually check clipped bits of data 
```{r}
# co2 (xlim=c(0.4247222, 0.4332870) corresponds to only MD1 first file)
par(mfrow=c(2,1),mar=c(4,4,1,1))
with(data_clean,plot(fDOY,CO2_dry,ylim=c(400,1300), xlim=c(0.4247222, 0.4332870)))
with(data_merge,plot(fDOY,CO2_dry,ylim=c(400,1300), xlim=c(0.4247222, 0.4332870)))

# ch4 
par(mfrow=c(2,1),mar=c(4,4,1,1))
with(data_clean,plot(fDOY,CH4_dry,ylim=c(1, 10), xlim=c(0.4247222, 0.4332870)))
with(data_merge,plot(fDOY,CH4_dry,ylim=c(1,10), xlim=c(0.4247222, 0.4332870)))

# weird august random check - MD 61
par(mfrow=c(2,1),mar=c(4,4,1,1))
with(data_clean3,plot(fDOY,CO2_dry,ylim=c(400,1700), xlim=c(219.5918, 219.5990)))
with(data_merge3,plot(fDOY,CO2_dry,ylim=c(400,1700), xlim=c(219.5918, 219.5990)))

# # ch4 
par(mfrow=c(2,1),mar=c(4,4,1,1))
with(data_clean3,plot(fDOY,CH4_dry,ylim=c(1, 25), xlim=c(219.5918, 219.5990)))
with(data_merge3,plot(fDOY,CH4_dry,ylim=c(1,25), xlim=c(219.5918, 219.5990)))
```

### TO DO - QC checking for all of the mounds 
* for each unique sample, determine the x bounds (min and max fDOY)
* then plot co2 and ch4 with these limits 
* visually check each mound to make sure correct portion is being clipped 

### Convert ppm to moles for CH4 and CO2
```{r chamber volumes and convert}
# use known volume of chamber (in m3) to convert ppm of gas to liters of gas
# small, h = 5.7 cm, d = 7.8 cm
sm <- pi*((7.8/2)^2)*5.7

# medium, h = 10.8 cm, d = 10.3 cm
med <- pi*((10.3/2)^2)*10.8

# large, h = 16.1 cm, d = 15.3 cm
lg <- pi*((15.3/2)^2)*16.1

# mini, v = 119 ml
mini <- 119

volume_cm3 <- c(sm, med, lg, mini)
chamber <- c("sm", "med", "lg", "mini")
chamber_vols <- data.frame(volume_cm3, chamber)

# chamber volumes are in cm3 - convert to m3 for the next step 
chamber_vols$volume_m3 <- chamber_vols$volume_cm3*0.000001

data_merge <- merge(data_merge, chamber_vols, by = c("chamber")) 
data_merge <- data_merge %>% relocate(chamber, .after = chamber_SA_cm2)

data_merge2 <- merge(data_merge2, chamber_vols, by = c("chamber")) 
data_merge2 <- data_merge2 %>% relocate(chamber, .after = chamber_SA_cm2)

data_merge3 <- merge(data_merge3, chamber_vols, by = c("chamber")) 
data_merge3 <- data_merge3 %>% relocate(chamber, .after = chamber_SA_cm2)

# also convert SA chamber to be in m2 (currently in cm2)
data_merge$chamber_SA_m2 <- data_merge$chamber_SA_cm2 * 0.0001
data_merge <- data_merge %>% relocate(chamber_SA_m2, .after = chamber_SA_cm2)

data_merge2$chamber_SA_m2 <- data_merge2$chamber_SA_cm2 * 0.0001
data_merge2 <- data_merge2 %>% relocate(chamber_SA_m2, .after = chamber_SA_cm2)

data_merge3$chamber_SA_m2 <- data_merge3$chamber_SA_cm2 * 0.0001
data_merge3 <- data_merge3 %>% relocate(chamber_SA_m2, .after = chamber_SA_cm2)

# CH4
data_merge$CH4_dry_L=
  # parts CH4 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge$CH4_dry/1000000) * data_merge$volume_m3 * 1000 
data_merge2$CH4_dry_L=
  # parts CH4 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge2$CH4_dry/1000000) * data_merge2$volume_m3 * 1000 
data_merge3$CH4_dry_L=
  # parts CH4 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge3$CH4_dry/1000000) * data_merge3$volume_m3 * 1000 

# CO2
data_merge$CO2_dry_L=
  # parts CO2 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge$CO2_dry/1000000) * data_merge$volume_m3 * 1000 
data_merge2$CO2_dry_L=
  # parts CO2 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge2$CO2_dry/1000000) * data_merge2$volume_m3 * 1000
data_merge3$CO2_dry_L=
  # parts CO2 per million parts air * volume of air in chamber (m3) * 1000 L per m3
  (data_merge3$CO2_dry/1000000) * data_merge3$volume_m3 * 1000 

# Use ideal gas law to calculate umol of CH4 or mmol of CO2
# CH4
data_merge$CH4_dry_umol=
  # (atm pressure * L CH4) / (R in L*atm/?K*mol * ?K temp) * 10^6 umol/mol
  ((1*data_merge$CH4_dry_L)/(0.08206*(data_merge$avg_respT+273)))*10^6
data_merge2$CH4_dry_umol=
  # (atm pressure * L CH4) / (R in L*atm/?K*mol * ?K temp) * 10^6 umol/mol
  ((1*data_merge2$CH4_dry_L)/(0.08206*(data_merge2$avg_respT+273)))*10^6
data_merge3$CH4_dry_umol=
  # (atm pressure * L CH4) / (R in L*atm/?K*mol * ?K temp) * 10^6 umol/mol
  ((1*data_merge3$CH4_dry_L)/(0.08206*(data_merge3$avg_respT+273)))*10^6

# CO2
data_merge$CO2_dry_mmol=
  # (atm pressure * L CO2) / (R in L*atm/?K*mol * ?K temp) * 10^3 mmol/mol
  ((1*data_merge$CO2_dry_L)/(0.08206*(data_merge$avg_respT+273)))*10^3
data_merge2$CO2_dry_mmol=
  # (atm pressure * L CO2) / (R in L*atm/?K*mol * ?K temp) * 10^3 mmol/mol
  ((1*data_merge2$CO2_dry_L)/(0.08206*(data_merge2$avg_respT+273)))*10^3
data_merge3$CO2_dry_mmol=
  # (atm pressure * L CO2) / (R in L*atm/?K*mol * ?K temp) * 10^3 mmol/mol
  ((1*data_merge3$CO2_dry_L)/(0.08206*(data_merge3$avg_respT+273)))*10^3
```
 
## Calculate chamber fluxes
```{r chamber fluxes}
## Identify start of fluxes
Time_start <- meta_TMI[,c("fDOY_start","sample")]
Time_start2 <- meta_TMI2[,c("fDOY_start","sample")]
Time_start3 <- meta_TMI3[,c("fDOY_start","sample")]
flux.times=unique(Time_start$fDOY_start)
flux.times2=unique(Time_start2$fDOY_start)
flux.times3=unique(Time_start3$fDOY_start)

# Create new dataframes to hold final fluxes 
fluxes.CH4_may=data.frame(matrix(NA,nrow=length(flux.times)))
fluxes.CO2_may=data.frame(matrix(NA,nrow=length(flux.times)))
fluxes.CH4_nov=data.frame(matrix(NA,nrow=length(flux.times2)))
fluxes.CO2_nov=data.frame(matrix(NA,nrow=length(flux.times2)))
fluxes.CH4_aug=data.frame(matrix(NA,nrow=length(flux.times3)))
fluxes.CO2_aug=data.frame(matrix(NA,nrow=length(flux.times3)))

## Add named columns
# CH4
fluxes.CH4_may$Time_start=flux.times
fluxes.CH4_may$flux.CH4=0
fluxes.CH4_may$R2.CH4=0
fluxes.CH4_may$p.CH4=0

fluxes.CH4_nov$Time_start=flux.times2
fluxes.CH4_nov$flux.CH4=0
fluxes.CH4_nov$R2.CH4=0
fluxes.CH4_nov$p.CH4=0

fluxes.CH4_aug$Time_start=flux.times3
fluxes.CH4_aug$flux.CH4=0
fluxes.CH4_aug$R2.CH4=0
fluxes.CH4_aug$p.CH4=0

# CO2
fluxes.CO2_may$Time_start=flux.times
fluxes.CO2_may$flux.CO2=0
fluxes.CO2_may$R2.CO2=0
fluxes.CO2_may$p.CO2=0

fluxes.CO2_nov$Time_start=flux.times2
fluxes.CO2_nov$flux.CO2=0
fluxes.CO2_nov$R2.CO2=0
fluxes.CO2_nov$p.CO2=0

fluxes.CO2_aug$Time_start=flux.times3
fluxes.CO2_aug$flux.CO2=0
fluxes.CO2_aug$R2.CO2=0
fluxes.CO2_aug$p.CO2=0

## Remove initial empty column
fluxes.CO2_may=fluxes.CO2_may[,-1]
fluxes.CH4_may=fluxes.CH4_may[,-1]
fluxes.CO2_nov=fluxes.CO2_nov[,-1]
fluxes.CH4_nov=fluxes.CH4_nov[,-1]
fluxes.CO2_aug=fluxes.CO2_aug[,-1]
fluxes.CH4_aug=fluxes.CH4_aug[,-1]

## For each start time (may 2022)
for (i in flux.times) {
  ## CH4 ##
  # Subset data for one chamber measurement
  temp1=subset(data_merge,fDOY_start==i)
  # Set corresponding row of output table
  j=which(flux.times==i)
  # Determine if start time has a CH4 flux
  if (nrow(temp1)>0) {
    # If so:  
    # Calulate flux in umol/day using linear regression
    mod=with(temp1,lm(CH4_dry_umol~fDOY))
    # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
    # flux rate
    fluxes.CH4_may$flux.CH4[j]=coef(mod)[2]/temp1$chamber_SA_m2 # chamber surface area
    # R2 of slope
    fluxes.CH4_may$R2.CH4[j]=summary(mod)$r.squared
    # p-value of slope
    fluxes.CH4_may$p.CH4[j]=summary(mod)$coefficients[2,4]
    # If not:
    # Fill rows of table with NA    
  } else {
    fluxes.CH4_may$flux.CH4[j]=NA
    fluxes.CH4_may$R2.CH4[j]=NA
    fluxes.CH4_may$p.CH4[j]=NA
  }
  ## CO2 ##
  # Subset data for one chamber measurement
  temp2=subset(data_merge,fDOY_start==i)
  # Calulate flux in mmol/day using linear regression
  mod=with(temp2,lm(CO2_dry_mmol~fDOY))
  # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
  # flux rate
  fluxes.CO2_may$flux.CO2[j]=coef(mod)[2]/temp2$chamber_SA_m2 # chamber surface area
  # R2 of slope
  fluxes.CO2_may$R2.CO2[j]=summary(mod)$r.squared
  # p-value of slope
  fluxes.CO2_may$p.CO2[j]=summary(mod)$coefficients[2,4]
}

## For each start time (nov 2022)
for (i in flux.times2) {
  ## CH4 ##
  # Subset data for one chamber measurement
  temp1=subset(data_merge2,fDOY_start==i)
  # Set corresponding row of output table
  j=which(flux.times2==i)
  # Determine if start time has a CH4 flux
  if (nrow(temp1)>0) {
    # If so:  
    # Calulate flux in umol/day using linear regression
    mod=with(temp1,lm(CH4_dry_umol~fDOY))
    # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
    # flux rate
    fluxes.CH4_nov$flux.CH4[j]=coef(mod)[2]/temp1$chamber_SA_m2 # chamber surface area
    # R2 of slope
    fluxes.CH4_nov$R2.CH4[j]=summary(mod)$r.squared
    # p-value of slope
    fluxes.CH4_nov$p.CH4[j]=summary(mod)$coefficients[2,4]
    # If not:
    # Fill rows of table with NA    
  } else {
    fluxes.CH4_nov$flux.CH4[j]=NA
    fluxes.CH4_nov$R2.CH4[j]=NA
    fluxes.CH4_nov$p.CH4[j]=NA
  }
  ## CO2 ##
  # Subset data for one chamber measurement
  temp2=subset(data_merge2,fDOY_start==i)
  # Calulate flux in mmol/day using linear regression
  mod=with(temp2,lm(CO2_dry_mmol~fDOY))
  # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
  # flux rate
  fluxes.CO2_nov$flux.CO2[j]=coef(mod)[2]/temp2$chamber_SA_m2 # chamber surface area
  # R2 of slope
  fluxes.CO2_nov$R2.CO2[j]=summary(mod)$r.squared
  # p-value of slope
  fluxes.CO2_nov$p.CO2[j]=summary(mod)$coefficients[2,4]
}

## For each start time (aug 2023)
for (i in flux.times3) {
  ## CH4 ##
  # Subset data for one chamber measurement
  temp1=subset(data_merge3,fDOY_start==i)
  # Set corresponding row of output table
  j=which(flux.times3==i)
  # Determine if start time has a CH4 flux
  if (nrow(temp1)>0) {
    # If so:  
    # Calulate flux in umol/day using linear regression
    mod=with(temp1,lm(CH4_dry_umol~fDOY))
    # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
    # flux rate
    fluxes.CH4_aug$flux.CH4[j]=coef(mod)[2]/temp1$chamber_SA_m2 # chamber surface area
    # R2 of slope
    fluxes.CH4_aug$R2.CH4[j]=summary(mod)$r.squared
    # p-value of slope
    fluxes.CH4_aug$p.CH4[j]=summary(mod)$coefficients[2,4]
    # If not:
    # Fill rows of table with NA    
  } else {
    fluxes.CH4_aug$flux.CH4[j]=NA
    fluxes.CH4_aug$R2.CH4[j]=NA
    fluxes.CH4_aug$p.CH4[j]=NA
  }
  ## CO2 ##
  # Subset data for one chamber measurement
  temp2=subset(data_merge3,fDOY_start==i)
  # Calulate flux in mmol/day using linear regression
  mod=with(temp2,lm(CO2_dry_mmol~fDOY))
  # Save flux rate and R2 and p-value of slope in corresponding row of dataframe
  # flux rate
  fluxes.CO2_aug$flux.CO2[j]=coef(mod)[2]/temp2$chamber_SA_m2 # chamber surface area
  # R2 of slope
  fluxes.CO2_aug$R2.CO2[j]=summary(mod)$r.squared
  # p-value of slope
  fluxes.CO2_aug$p.CO2[j]=summary(mod)$coefficients[2,4]
}
```

## Finalize fluxes, merge back with metadata and extract
```{r}
# merge metadata back in to flux files 
# CO2
meta_TMI <- meta_TMI %>%
  dplyr::rename(Time_start = fDOY_start)
meta_TMI2 <- meta_TMI2 %>%
  dplyr::rename(Time_start = fDOY_start)
meta_TMI3 <- meta_TMI3 %>%
  dplyr::rename(Time_start = fDOY_start)

CO2_fluxfinal_may22 <- merge(fluxes.CO2_may, meta_TMI, by="Time_start")
CO2_fluxfinal_may22$campaign <- "may22"
CO2_fluxfinal_nov22 <- merge(fluxes.CO2_nov, meta_TMI2, by="Time_start")
CO2_fluxfinal_nov22$campaign <- "nov22"
CO2_fluxfinal_aug23 <- merge(fluxes.CO2_aug, meta_TMI3, by="Time_start")
CO2_fluxfinal_aug23$campaign <- "aug23"

CO2_fluxfinal_all <- rbind.fill(CO2_fluxfinal_may22, CO2_fluxfinal_nov22, CO2_fluxfinal_aug23)

# CH4
CH4_fluxfinal_may22 <- merge(fluxes.CH4_may, meta_TMI, by="Time_start")
CH4_fluxfinal_may22$campaign <- "may22"
CH4_fluxfinal_nov22 <- merge(fluxes.CH4_nov, meta_TMI2, by="Time_start")
CH4_fluxfinal_nov22$campaign <- "nov22"
CH4_fluxfinal_aug23 <- merge(fluxes.CH4_aug, meta_TMI3, by="Time_start")
CH4_fluxfinal_aug23$campaign <- "aug23"

CH4_fluxfinal_all <- rbind.fill(CH4_fluxfinal_may22, CH4_fluxfinal_nov22, CH4_fluxfinal_aug23)

## Export fluxes as .csv file
# CO2 (units are mmol/day)
# write.csv(CO2_fluxfinal_all,"/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CO2_fluxfinal_all.csv", row.names = FALSE)

# CH4 (units are umol/day)
# write.csv(CH4_fluxfinal_all,"/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CH4_fluxfinal_all.csv", row.names = FALSE)
```
