---
title: "2.prelim_fluxviz"
author: "abbey yatsko"
date: "26/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(viridis)
library(tidyr)
```

read in data 
```{r data load}
CO2_fluxfinal <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CO2_fluxfinal_all.csv")
CH4_fluxfinal<- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CH4_fluxfinal_all.csv")
```

data clean and reorganize
```{r}
# merge in species data to flux data 
species <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/Abbey termites .xlsx - Sheet1.csv")
species <- species[c("X", "ID_cleaned")]

# merge sp and CH4flux_final, CO2_fluxfinal
colnames(species)[1] = "sample"

CH4_fluxfinal <- merge(CH4_fluxfinal, species, by = c("sample")) 
CO2_fluxfinal <- merge(CO2_fluxfinal, species, by = c("sample")) 

# new column for methane:co2
# first need to convert CO2 from mmol to umol (in order to match CH4 in units of umol)
# (conversion factor 1000umol = 1mmol)
CO2_fluxfinal$flux.CO2_umol <- CO2_fluxfinal$flux.CO2*1000
CO2_fluxfinal <- CO2_fluxfinal %>% relocate(flux.CO2_umol, .after = flux.CO2)

CH4_fluxfinal$ch4_co2 <-  CH4_fluxfinal$flux.CH4/CO2_fluxfinal$flux.CO2
CH4_fluxfinal <- CH4_fluxfinal %>% relocate(ch4_co2, .after = flux.CH4)

# new column to add arboreal where 1 = arboreal, 0 = not arboreal
CO2_fluxfinal$arboreal[CO2_fluxfinal$chamber == "mini"] <- 1
CO2_fluxfinal$arboreal[!CO2_fluxfinal$chamber == "mini"] <- 0

CH4_fluxfinal$arboreal[CH4_fluxfinal$chamber == "mini"] <- 1
CH4_fluxfinal$arboreal[!CH4_fluxfinal$chamber == "mini"] <- 0

# subset df to create just methane from soils and just mounds (for later on. just to have df prepped)
soils <- CH4_fluxfinal[CH4_fluxfinal$flux_source == 's',]
mounds <- CH4_fluxfinal[CH4_fluxfinal$flux_source == 'm',]

# arboreal only 
arboreal <- mounds[mounds$arboreal == 1,]

# omit arboreal from mound measurements 
mounds <- mounds[mounds$arboreal == 0,]

# remove unidentified mounds 
mounds[mounds==""] <- NA
mounds_ID <- mounds[!is.na(mounds$ID_cleaned),]

# remove duplicate ID columns 
mounds_ID <- subset( mounds_ID, select = -c(ID_cleaned.x, ID_cleaned.y) )

# write.csv(mounds, "/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/mounds.csv")
```

methane flux for individual mounds, comparing seasons, by species
```{r flux}
# all mounds
a1 <- ggplot(data = mounds_ID, aes(x = sample, y = flux.CH4, fill = campaign)) + 
  geom_boxplot()+ 
  facet_wrap(~ID_cleaned, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,18000)+ # removes outliers, makes it easier to see comparisons
  ylab("CH4 flux umol/d/m2")+
  labs(title = "mound")

# all soils
# a2 <- ggplot(data = soils, aes(x = as.factor(sample), y = flux.CH4, fill = ID_cleaned)) + 
#   geom_boxplot()+ 
#   geom_jitter()+
#   theme_classic()+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   ylab("CH4 flux umol/d/m2")+
#   labs(title = "soil")
# ggarrange(a1, a2, ncol=1)

# individual mounds with CH4:CO2
a3 <- ggplot(data = mounds_ID, aes(x = as.factor(sample), y = ch4_co2, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_wrap(~campaign, scales = "free_x")+
  ylim(0,50)+ # for outliers
  labs(title = "mound")
  
# a4 <- ggplot(data = soils, aes(x = as.factor(sample), y = ch4_co2, fill = ID_cleaned)) + 
#   geom_boxplot()+ 
#   theme_classic()+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   ylim(0,2.5)+ # for outliers
#   labs(title = "soil")
# ggarrange(a3, a4, ncol=1)
```

show significant mound methane flux to hardly any soil flux 
```{r by soil/mound}
b <- ggplot(data = CH4_fluxfinal, aes(x = as.factor(flux_source), y = flux.CH4)) + 
  geom_boxplot()+ 
  geom_jitter()+
  theme_light()
```

mound and soil methane flux by position
```{r by position}
# mounds 
c1 <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = flux.CH4, fill = position)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  scale_fill_brewer("Blues")+
  theme_light()
c2 <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = ch4_co2, fill = position)) + 
  geom_boxplot()+
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4:CO2")+
  scale_fill_brewer("Blues")+
  ylim(0,50)+ # for outliers
  theme_light()

ggarrange(c1, c2, ncol = 1)

# soils (shows decreasing as you move from the mound, but all together insignificant flux)
c3 <- ggplot(data = soils, aes(x = as.factor(position), y = flux.CH4, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  scale_fill_brewer("Blues")+
  theme_light()
```

methane flux for mounds based on cardinal direction of measurement
```{r by direction}
# all mounds, across different directions
d1 <- ggplot(data = mounds_ID, aes(x = as.factor(directon), y = flux.CH4)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  theme_light()
d2 <- ggplot(data = mounds_ID, aes(x = as.factor(directon), y = ch4_co2)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  theme_light()

ggarrange(d1, d2, ncol = 1)

# plus by species, across different directions
d3 <- ggplot(data = mounds_ID, aes(x = ID_cleaned, y = flux.CH4, fill = as.factor(directon))) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  theme_light()
d4 <- ggplot(data = mounds_ID, aes(x = ID_cleaned, y = flux.CH4, fill = as.factor(directon))) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  theme_light()

ggarrange(d3, d4, ncol = 1)
```

average methane flux by species, contrasted between seasons
```{r by species}
# overall methane flux
e <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = flux.CH4, fill = campaign)) + 
  geom_boxplot()+ 
  scale_fill_brewer("Blues")+
  ylab("CH4 flux umol/d/m2")+
  theme_light()

# co2:ch4 ratio
 f <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = ch4_co2, fill = campaign)) + 
  geom_boxplot()+ 
  theme_light()+
  scale_fill_brewer("Blues")+
  scale_y_continuous(limits = c(-1, 50))

ggarrange(e, f, ncol = 1)
```

methane flux by mound average for different parts of the day 
```{r}
# merge in real time data from GPS records of mound being tagged (this is an overall mound estimate for the may campaign mounds)
maytime <- read.csv("/Users/abbeyyatsko/Downloads/mound_gps (1).csv")
maytime <- maytime[c("name", "realtime")]
names(maytime)[1] <- 'sample'
maymounds <- mounds[mounds$campaign == "may22",]
novmounds <- mounds[mounds$campaign == "nov22",]

maymounds <- merge(maymounds, maytime, by = c("sample")) 
maymounds <- maymounds %>% relocate(realtime, .after = Time_start)

novtime <- read.csv("/Users/abbeyyatsko/Downloads/novtimes.csv")
names(novtime)[2] <- 'realtime'
novmounds <- merge(novmounds, novtime, by = c("sample")) 
novmounds <- novmounds %>% relocate(realtime, .after = Time_start)

# remove duplicate ID columns 
maymounds <- subset( maymounds, select = -c(ID_cleaned.x, ID_cleaned.y) )
novmounds <- subset( novmounds, select = -c(ID_cleaned.x, ID_cleaned.y) )

# plot for flux by time of day - may
g1 <- ggplot(data = maymounds, aes(x = realtime, y = flux.CH4, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  scale_fill_brewer("Blues")+
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  theme_light()

# plot for flux by time of day - nov
g2 <- ggplot(data = novmounds, aes(x = realtime, y = flux.CH4, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  scale_fill_brewer("Blues")+
  scale_x_discrete(limits=c("8:36:45", "8:52:40", "9:13:40", "9:30:10", "9:36:40", "10:07:00", "10:24:30", 
                            "10:58:00", "11:40:40", "11:43:30", "12:02:15", "12:16:00", "13:32:15", "13:50:30", 
                            "13:52:00", "14:08:30", "14:22:30", "14:31:00", "14:47:00", "14:51:00", "14:57:15",
                            "15:34:15", "15:55:40", "16:23:00"), guide = guide_axis(angle = 90)) +
  theme_light()

# # scatter plot for flux by time of day (for soils)
# g3 <- ggplot(data = soils, aes(x = as.factor(realtime), y = flux.CH4, fill = ID_cleaned)) + 
#   geom_boxplot()+ 
#   ylab("CH4 flux umol/d/m2")+
#   scale_x_discrete(guide = guide_axis(angle = 90)) +
#   theme_light()

ggarrange(g1, g2, ncol = 1)

# using radial graph for mounds
# g4 <- ggplot(data = moundsCO2, aes(x = as.factor(realtime), y = flux.CH4, fill = ID_cleaned)) + 
#   geom_boxplot()+ 
#   ylab("CH4 flux umol/d/m2")+
#   scale_y_continuous(limits = c(-1, 15000))+
#   theme_light()+
#   coord_polar() 
```

methane flux by temperature 
```{r}
# scatter plot - methane 
h1 <- ggplot(data = mounds_ID, aes(x = avg_respT, y = flux.CH4, color = campaign)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  ylab("CH4 flux umol/d/m2")+
  facet_wrap(~ID_cleaned, scales = "free_x")+
  scale_color_brewer("Reds") +
  theme_linedraw()

# average temperature by mound via campaign
mounds_ID$mound_campaign <- paste(mounds_ID$sample,mounds_ID$campaign, sep = "-")
x <- aggregate(mounds_ID$avg_respT, list(mounds_ID$mound_campaign), FUN=mean)
colnames(x)[1] = "mound_campaign"
colnames(x)[2] = "mound_temp"
mounds_ID <- merge(mounds_ID, x, by = c("mound_campaign")) 
str(mounds_ID)
mounds_ID$campaign <- as.factor(mounds_ID$campaign)

h2 <- ggplot(data = mounds_ID, aes(x = mound_temp, y = flux.CH4, fill = as.factor(sample))) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  facet_grid(campaign~ID_cleaned, scales = "free_x")+
  theme(legend.position="none")+
  scale_fill_viridis(discrete = TRUE, option = "D")+
  ylim(0,16000) # for outliers

```

scaled mound flux from calculated SA (come back to this)
```{r}
# find average flux by mound, calculating seperately remeasured mounds  
# group by mound_campaign
y <- mounds_ID %>% group_by(mound_campaign) %>% 
  summarise(mean_flux=mean(flux.CH4))
y <- separate(y, col=mound_campaign, into=c('sample', 'campaign'), sep='-')

# merge with mound_SA
mound_SA <- read.csv("/Users/abbeyyatsko/Downloads/mound SA - Sheet1 (1).csv")
mound_SA <- mound_SA[, c('sample', 'working_sa_m2')]
mound_SA <- mound_SA[!is.na(mound_SA$working_sa_m2),]
z <- merge(y, mound_SA, by = "sample" )
z$SA_scaled_flux <- z$mean_flux * z$working_sa_m2
z <- merge(z, species, by = "sample" )

# graph mound flux by mound size
ggplot(data = z, aes(x = working_sa_m2, y = SA_scaled_flux)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  facet_wrap(~ID_cleaned, scales = "free_x")+
  theme_classic()+
  ylab("CH4 flux umol/d/m2 scaled to mound size") +
  xlab("mound surface area")
```

calculate average flux per species 
```{r}
# average fluxes - may 
# remove NAs 
maymounds_ID <- maymounds[!is.na(maymounds$ID_cleaned),]

may_avg <- maymounds_ID %>% group_by(ID_cleaned) %>% 
  summarise(mean_flux=mean(flux.CH4),
            .groups = 'drop')

# average fluxes - nov 
nov_avg <- novmounds %>% group_by(ID_cleaned) %>% 
  summarise(mean_flux=mean(flux.CH4),
            .groups = 'drop')

# to convert umol/m2*d to ug/m2*d:
# avgflux_ami <- avgflux_ami*(1/1000000)*(16.04)*(1000000) # molecular mass conversion to g, then conversion back to ug for comparison 
# avgflux_nasute <- avgflux_nasute*(1/1000000)*(16.04)*(1000000) 
# avgflux_copto <- avgflux_copto*(1/1000000)*(16.04)*(1000000) 
# final units for average fluxes are now in ug CH4/m2*d
```












# methane vs. co2 regression 
```{r}
# new df with sample, ch4 and co2 
reg <- merge(CH4_fluxfinal, CO2_fluxfinal, by = c("Time_start")) 
reg <- reg[, c("Time_start", "flux.CH4", "flux.CO2", "flux.CO2_umol", "flux_source.x", "ID_cleaned.x", "avg_respT.x" )]
str(reg)
reg_mounds <- reg[reg$flux_source.x == 'm',]
reg_soils <- reg[reg$flux_source.x == 's',]

i1 <- ggplot(data = reg, aes(x = flux.CH4, y = flux.CO2_umol, color = ID_cleaned.x)) + 
  geom_point(alpha=0.5)+ 
  xlab("CH4 flux umol/d/m2")+
  ylab("CO2 flux umol/d/m2")+
  geom_smooth(method = 'lm')+
  labs(title = "mound+soil")+
  theme_light()
i2 <- ggplot(data = reg_mounds, aes(x =log(flux.CO2_umol) , y =log(flux.CH4), color = ID_cleaned.x)) +
  geom_point(alpha=0.5)+ 
  ylab("CH4 flux umol/d/m2")+
  xlab("CO2 flux umol/d/m2")+
  geom_smooth(method = 'lm')+
  labs(title = "mound")+
  theme_light()
i3 <- ggplot(data = reg_soils, aes(x = flux.CH4, y = flux.CO2_umol, color = ID_cleaned.x)) + 
  geom_point(alpha=0.5)+ 
  xlab("CH4 flux umol/d/m2")+
  ylab("CO2 flux umol/d/m2")+
  geom_smooth(method = 'lm')+
  labs(title = "soil")+
  theme_light()
ggarrange(i1, i2, i3)
```
