---
title: "3.choosefor_methanotrophs"
author: "abbey yatsko"
date: "28/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The goal for this document is to sort out mound methane flux data and choose samples for sequencing 

There are two main research questions at this time: 
  1. How does the mound methanotroph community composition change across different termite species?
  2. For termite mounds with high inter-mound variation in methane flux, what is variation in inter-mound methanotroph community c       composition?

### set workspace, read data 
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)

mounds <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/mounds.csv")
mounds <- mounds[,-1]

# remove entries without species ID 
mounds[mounds==""] <- NA
mounds <- mounds[!is.na(mounds$ID_cleaned),]

# remove mounds that have fewer than 4 measurement positions per mound 
measure_number <- data.frame(table(mounds$sample))
measure_number <- subset(measure_number, Freq>3)
measure_number <- as.vector(measure_number$Var1)
mounds <- mounds[mounds$sample %in% measure_number,]
```

### To address question 1: 
first, is mound flux different by species?
```{r}
# one way ANOVA
aov.1 <- aov(flux.CH4 ~ ID_cleaned, data = mounds) # note experimental switch out of CH4 flux with CH4:CO2
summary(aov.1)
TukeyHSD(aov.1)

# assumptions
par(mfrow=c(2,2))
plot(aov.1)
par(mfrow=c(1,1))
```

within each mound sample, is mound flux different based on position? 
```{r}
mound1 <- mounds[mounds$sample == c('MD1'),]
aov.2 <- aov(flux.CH4 ~ position, data = mound1) # note experimental switch out of CH4 flux with CH4:CO2
summary(aov.2)

# find a way to loop through each mound number and calculate anova table? 
```

###  To address question 1
Contrast big and little methane emitters
  * target top 4 and bottom 4 overall mound fluxes (pooled from all positions) for 3 species 
  * 8 mounds * 3 species * 4-5 positions 
```{r}
avgflux <- aggregate(mounds$flux.CH4, list(mounds$sample), FUN=mean)
y <- mounds[, c("sample",  "ID_cleaned")]

avgflux <- left_join(avgflux, y, by = c("Group.1"="sample"))
avgflux <- unique(avgflux)

# average methane flux by species
ggplot(data = avgflux, aes(x = as.factor(ID_cleaned), y = x)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  theme_light()

# anova on mound averages between species
aov.3 <- aov(x ~ ID_cleaned, data = avgflux)
summary(aov.3)
TukeyHSD(aov.3)

# pick out top 4 highest methane fluxes by species 
# (there is overlap between high and low of amitermes therefore shift from 5 to 4)
top <- avgflux %>%                                     
  arrange(desc(x)) %>% 
  group_by(ID_cleaned) %>%
  slice(1:4)

top$rank <- "top"
top    

# pick out bottom 4 lowest methane fluxes by species
bottom <- avgflux %>%                                     
  arrange(x) %>% 
  group_by(ID_cleaned) %>%
  slice(1:4)

bottom$rank <- "bottom"
bottom   

# remove entries without species info, clean up df
q1 <- rbind(top, bottom)
q1 <- q1 %>%
  rename(sample = Group.1, 
         avg_flux = x)

# now take these IDs and show boxplots to see distribution of fluxes (not just average) 
# join df
q1_all <- left_join(q1, mounds, by = "sample")

# plot for all
ggplot(data = q1_all, aes(x = sample, y = flux.CH4, color =  as.factor(rank), fill = ID_cleaned.x)) + 
  geom_boxplot()+ 
  geom_jitter()+
  ylab("CH4 flux umol/d/m2")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~ as.factor(ID_cleaned.x), 
             scales = "free_x",
             space = "free_x",
             switch = "x")+ 
  scale_fill_brewer(palette="YlOrRd") 

# bar graph grouped by species - colored by top and bottom 
ggplot(q1, aes(x = reorder(as.factor(sample), -avg_flux), y = avg_flux, fill = rank)) + 
  geom_bar(stat='identity') +
  facet_grid(~ as.factor(ID_cleaned), 
             scales = "free_x",
             space = "free_x",
             switch = "x")+ 
  ylab("CH4 flux umol/d/m2")+
  xlab("")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# plot for high values
q1H <- q1_all[q1_all$rank == c('top'),]
q1Hplot <- ggplot(data = q1H, aes(x = sample, y = flux.CH4, fill = ID_cleaned.x)) + 
  geom_boxplot()+ 
  geom_jitter()+
  ylab("CH4 flux umol/d/m2")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")+
  facet_grid(~ as.factor(ID_cleaned.x), 
             scales = "free_x",
             space = "free_x",
             switch = "x")+ 
  scale_fill_brewer(palette="YlOrRd")+
  ggtitle("mounds w highest average flux")

# plot for low values 
q1L <- q1_all[q1_all$rank == c('bottom'),]
q1Lplot <- ggplot(data = q1L, aes(x = sample, y = flux.CH4, fill = ID_cleaned.x)) + 
  geom_boxplot()+ 
  geom_jitter()+
  ylab("CH4 flux umol/d/m2")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")+
  facet_grid(~ as.factor(ID_cleaned.x), 
             scales = "free_x",
             space = "free_x",
             switch = "x")+ 
  scale_fill_brewer(palette="YlOrRd")+ 
  ggtitle("mounds w lowest average flux")

# overall sampling for Q1
ggarrange(q1Hplot, q1Lplot, ncol = 1)

# export .csv with q1 mounds 
# write.csv(q1, "/Users/abbeyyatsko/Downloads/q1.csv")
```

### To address question 2: 
explore intramound methane flux variation
target highest within-mound variation in methane flux and sequence each positon individually 
  * top 5 most variable mounds * 3 species * 4-5 positions =  60-75 samples
```{r}
# calculate min and max flux for each sample 
data_min <- mounds %>%                                     
  group_by(sample) %>%
  summarise_at(vars(flux.CH4),
               list(min = min))
data_max <- mounds %>%                                     
  group_by(sample) %>%
  summarise_at(vars(flux.CH4),
               list(max = max))

minmax <- merge(data_max, data_min, by="sample")
minmax$range <- minmax$max - minmax$min 

# merge in species info 
minmax <- left_join(minmax, y, by = "sample")
minmax <- unique(minmax)

# range in intermound methane production by sample 
ggplot(data = minmax, aes(x = as.factor(ID_cleaned), y = range)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  theme_light()

# choose 5 most variable mounds (greatest range)
q2 <- minmax %>%                                     
  arrange(desc(range)) %>% 
  group_by(ID_cleaned) %>%
  slice(1:5)

# boxplot of individual mounds
q2_samps <- as.vector(q2$sample)
q2_samps <- mounds[ mounds$sample %in% q2_samps, ]
unique(q2_samps$sample)

ggplot(data = q2_samps, aes(x = as.factor(sample), y = flux.CH4, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  geom_jitter()+
  facet_grid(~ as.factor(ID_cleaned), 
             scales = "free_x",
             space = "free_x",
             switch = "x")+ 
  ylab("CH4 flux umol/d/m2")+
  xlab("")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")

# see the overlap of samples in q1 and q2 
matches <- inner_join(q1, q2, 
                       c("sample" = "sample"))
# export .csv with q2 mounds 
# write.csv(q2, "/Users/abbeyyatsko/Downloads/q2.csv")
```

full sample lists 
```{r}
# q1: choose 3 samples on each mound closest to average flux
# create unique extract_ID that includes mound number, tag, and position of measurement
q1_samps <- q1_all
q1_samps$tag <- as.factor(q1_samps$tag)
q1_samps$extract_ID <- paste(q1_samps$sample,"-",q1_samps$tag, "-", q1_samps$position, sep = "")
q1_samps <- q1_samps %>% relocate(extract_ID, .after = sample)

# choose 3 samples per mound that are closest to the average 
# make column for difference from median in order to sort out easier 
q1_samps$avg_dif <- abs(q1_samps$avg_flux - q1_samps$flux.CH4)
q1_samps <- q1_samps %>% relocate(avg_dif, .after = avg_flux)

# choose the largest average difference grouped by sample 
q1_samps %>% group_by(sample) %>% top_n(1, avg_dif) -> a
a <- as.vector(a$extract_ID)
q1_samps <- q1_samps[!q1_samps$extract_ID %in% a,]

# subset out by species 
q1_samps_ami <- subset(q1_samps, ID_cleaned.x == "Ami")
q1_samps_copto <- subset(q1_samps, ID_cleaned.x == "Copto")
q1_samps_nasuti <- subset(q1_samps, ID_cleaned.x == "Nasuti")

# get rid of highest again for copto and nasuti (they have 5 measurements per mound usually, need to rid of an additional)
q1_samps_copto %>% group_by(sample) %>% top_n(1, avg_dif) -> copto_rm
copto_rm <- as.vector(copto_rm$extract_ID)
q1_samps_copto <- q1_samps_copto[!q1_samps_copto$extract_ID %in% copto_rm,]

q1_samps_nasuti %>% group_by(sample) %>% top_n(1, avg_dif) -> nasuti_rm
nasuti_rm <- as.vector(nasuti_rm$extract_ID)
q1_samps_nasuti <- q1_samps_nasuti[!q1_samps_nasuti$extract_ID %in% nasuti_rm,]

# bring species back together 
q1_samps <- rbind(q1_samps_ami, q1_samps_copto, q1_samps_nasuti)

# export as .csv 
# write.csv(q1_samps, "/Users/abbeyyatsko/Downloads/q1_samps.csv", row.names = FALSE)

# q1 minus overlaps: remove samples that are already being accounted for in q2 
# every position for mounds in q2 are being extracted
matches <- as.vector(matches$sample)
q1_samps_no_overlap <- q1_samps[!q1_samps$sample %in% matches,]
# write.csv(q1_samps_no_overlap, "/Users/abbeyyatsko/Downloads/q1_samps_no_overlap.csv", row.names = FALSE)

# q2: sequencing all individual samples within select mounds
dim(q2_samps)
# 68 samples (from 15 mounds)

# create unique extract_ID that includes mound number, tag, and position of measurement
q2_samps$tag <- as.factor(q2_samps$tag)
q2_samps$extract_ID <- paste(q2_samps$sample,"-",q2_samps$tag, "-", q2_samps$position, sep = "")
q2_samps <- q2_samps %>% relocate(extract_ID, .after = sample)

# export as .csv and check what overlap exists between q2 and q1
# write.csv(q2_samps, "/Users/abbeyyatsko/Downloads/q2_samps.csv", row.names = FALSE)

# total extractions: 115

```

