---
title: "3.choosefor_methanotrophs"
author: "abbey yatsko"
date: "28/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The goal for this document is to sort out mound methane flux data and choose samples for sequencing 

There are two main research questions at this time: 
  1. How does the mound methanotroph community composition change across different termite species?
  2. For termite mounds with high inter-mound variation in methane flux, what is variation in inter-mound methanotroph community c       composition?

### set workspace, read data 
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)

CH4_fluxfinal<- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CH4_fluxfinalTMI.csv")
species <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/Abbey termites .xlsx - Sheet1.csv")
species <- species[c("X", "ID_cleaned")]

# merge sp and CH4flux_final, CO2_fluxfinal
species <- species %>%
  rename(sample = X)
CH4_fluxfinal <- merge(CH4_fluxfinal, species, by = c("sample")) 

# subset out soils to have a df containing only mound information 
mounds <- CH4_fluxfinal[CH4_fluxfinal$flux_source == 'm',]
```

### To address question 1: 
first, is mound flux different by species?
```{r}
# one way ANOVA
aov.1 <- aov(flux.CH4 ~ ID_cleaned, data = mounds) # note experimental switch out of CH4 flux with CH4:CO2
summary(aov.1)
TukeyHSD(aov.1)

# assumptions
par(mfrow=c(2,2))
plot(aov.1)
par(mfrow=c(1,1))
```

within each mound sample, is mound flux different based on position? 
```{r}
mound1 <- mounds[mounds$sample == c('MD1'),]
aov.2 <- aov(flux.CH4 ~ position, data = mound1) # note experimental switch out of CH4 flux with CH4:CO2
summary(aov.2)

# find a way to loop through each mound number and calculate anova table? 
```

contrast big and little methane emitters
  * target top 5 and bottom 5 overall mound fluxes (pooled from all positions) for 3 species 
  * 10 mounds * 3 species = 30 samples 
```{r}
avgflux <- aggregate(mounds$flux.CH4, list(mounds$sample), FUN=mean)
y <- mounds[, c("sample",  "ID_cleaned")]

avgflux <- left_join(avgflux, y, by = c("Group.1"="sample"))
avgflux <- unique(avgflux)

# average methane flux by species
ggplot(data = avgflux, aes(x = as.factor(ID_cleaned), y = x)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  theme_light()

# anova on mound averages between species
aov.3 <- aov(x ~ ID_cleaned, data = avgflux)
summary(aov.3)
TukeyHSD(aov.3)

# pick out top 5 highest methane fluxes by species
top <- avgflux %>%                                     
  arrange(desc(x)) %>% 
  group_by(ID_cleaned) %>%
  slice(1:5)
top    

# pick out bottom 5 lowest methane fluxes by species
bottom <- avgflux %>%                                     
  arrange(x) %>% 
  group_by(ID_cleaned) %>%
  slice(1:5)
bottom   
```

### To address question 2: 
explore intermound methane flux variation
target highest within-mound variation in methane flux and sequence each positon individually 
  * top 5 most variable mounds * 3 species * 4-5 positions =  60-75 samples
```{r}
# calculate min and max flux for each sample 
data_min <- mounds %>%                                     
  group_by(sample) %>%
  summarise_at(vars(flux.CH4),
               list(min = min))
data_max <- mounds %>%                                     
  group_by(sample) %>%
  summarise_at(vars(flux.CH4),
               list(max = max))

minmax <- merge(data_max, data_min, by="sample")
minmax$range <- minmax$max - minmax$min 

# merge in species info 
minmax <- left_join(minmax, y, by = "sample")
minmax <- unique(minmax)

# range in intermound methane production by sample 
ggplot(data = minmax, aes(x = as.factor(ID_cleaned), y = range)) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  theme_light()

# choose 5 most variable mounds (greatest range)
variable <- minmax %>%                                     
  arrange(desc(range)) %>% 
  group_by(ID_cleaned) %>%
  slice(1:5)
variable  
```



