---
title: "7.flux_viz"
author: "abbey yatsko"
date: "26/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(plyr)
library(ggpubr)
#library(patchwork)
library(tidyr)
library(easystats)
library(ggeffects)
library(emmeans)
library(lme4)
library(lmerTest)
library(splines)
```

data load 
```{r}
all_fluxes_NT <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/all_fluxes_NT.csv")
p_fluxes_f <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/p_fluxes_f.csv")
all_fluxes_mound_avg <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/all_fluxes_mound_avg.csv")
all_fluxes_mound <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/all_fluxes_mound.csv")
all_fluxes_seasonal_avg <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/all_fluxes_resample_reps_avg.csv")

# supplement (?)
all_fluxes_daily <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/all_daily_fluxes.csv")
```

Q1: How does CH4 flux vary for different termite species, at the a) individual termite and b) mound level?
```{r}
# a) INDIVIDUAL LEVEL 
# TEF 
ggplot(all_fluxes_NT, aes(x = species, y = TEF_ug_g_h, fill = species))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter()+
  ylab("Termite EF (ug CH4/hr/g termite)")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()+
  theme(legend.position = "none")

# test for differences between species 
mod <- aov(TEF_ug_g_h ~ species, data = all_fluxes_NT)
summary(mod)
# determine which species are different
emmeans(mod, pairwise ~ species)

# b) MOUND LEVEL
ggplot(all_fluxes_mound_avg, aes(x = species_s, y = mean_mound_fluxCH4, fill = species_s)) + 
  geom_boxplot()+ 
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  theme_classic()+
  ylab("CH4 flux umol/d/m2")+
  xlab("")+
  scale_fill_brewer(palette = "Dark2")+
  theme(legend.position = "none")

# test for differences between species 
mod <- aov(mean_mound_fluxCH4 ~ species_s, data = all_fluxes_mound_avg)
summary(mod)
# determine which species are different
emmeans(mod, pairwise ~ species_s)
```

Q2: How does CH4 flux change for different mound properties: a) mound size and b) wall thickness
```{r}
# a) MOUND SIZE
# average mound size by species
ggplot(all_fluxes_mound_avg, aes(x = species_s, y = SA_m2, fill = species_s))+ # swap out basal_area_m2
  geom_boxplot()+
  theme_classic()+
  xlab("")+
  ylab("Mound surface area (m2)")+
  scale_fill_brewer(palette = "Dark2")+
  theme(legend.position = "none")

mod <- aov(SA_m2 ~ species_s, data = all_fluxes_mound_avg)
summary(mod)
emmeans(mod, pairwise ~ species_s)

# methane flux (whole mound) by mound size 
ggplot(all_fluxes_mound_avg, aes(x = SA_m2, y = mound_CH4flux, color = species_s)) + # swap out basal_area_m2, color = campaign
  geom_point()+ 
  geom_smooth(method = 'lm')+
  facet_wrap(~species_s, scales = "free")+ 
  scale_color_brewer(palette = "Dark2")+
  theme_classic()+
  ylab("Total mound CH4 flux (umol/d/mound)") +
  xlab("Mound surface area (m2)")+
  theme(legend.position = "none")

mod <- lm(mound_CH4flux ~ SA_m2 + species_s, data = all_fluxes_mound_avg)
summary(mod)
report(mod)
ggpredict(mod, terms = c("SA_m2", "species_s")) %>% plot(add.data = TRUE)

# b) WALL THICKNESS
wt <- all_fluxes_mound[!is.na(all_fluxes_mound$wall_thickness_mm), ]
wt$sample <- as.factor(wt$sample)

# to calculate average mound wall thickness
# wt_avg <- wt %>% group_by(ID_mound, sample, species_s) %>% 
#   summarise(mean_mound_fluxCH4=mean(flux.CH4),
#             mean_mound_fluxCO2=mean(flux.CO2_umol), 
#             mean_mound_fluxCH4_CO2=mean(CH4_CO2), 
#             campaign = campaign, 
#             temp = mean(avg_respT), 
#             precip_1mo = precip_1mo, 
#             wall_thickness_mm = mean(wall_thickness_mm))
# wt_mound_avg <- distinct(wt_avg)

# average wall thickness by species
ggplot(wt, aes(x = species_s, y = wall_thickness_mm, fill = species_s))+ # swap out basal_area_m2
  geom_boxplot()+
  theme_classic()+
  xlab("")+
  ylab("Mound wall thickness (mm)")+
  scale_fill_brewer(palette = "Dark2")+
  theme(legend.position = "none")

mod <- aov(wall_thickness_mm ~ species_s, data = wt)
summary(mod)
emmeans(mod, pairwise ~ species_s)

ggplot(wt, aes(x = wall_thickness_mm, y = flux.CH4, color = species_s)) +
  geom_point()+ 
  geom_smooth(method = "lm")+
  facet_wrap(~species_s, scales = "free")+ 
  theme_classic()+
  ylab("CH4 flux umol/d/m2")+
  xlab("Mound wall thickness (mm)")+
  ggtitle("Sample point mound wall thickness")+
  scale_color_brewer(palette = "Dark2")+
  theme(legend.position = "bottom")

mod <- lmer(flux.CH4 ~ wall_thickness_mm + species_s + (1 | sample), data = wt)
summary(mod)
Anova(mod)
report(mod)
ggpredict(mod, terms = c("wall_thickness_mm", "species_s")) %>% plot(add.data = TRUE)
```

Q3: How does CH4 flux change annually, and is this driven by environmental conditions (temperature and precipitation)?
```{r}
# a) seasonal comparison 
all_fluxes_seasonal_avg$campaign <- factor(all_fluxes_seasonal_avg$campaign, levels = c("may22", "aug23", "nov22", "feb24"))
all_fluxes_seasonal_avg$species_s <- as.factor(all_fluxes_seasonal_avg$species_s)

ggplot(all_fluxes_seasonal_avg, aes(x = campaign, y = mean_mound_fluxCH4, fill = campaign))+
  geom_boxplot()+
  ylab("CH4 flux umol/d/m2")+
  xlab("")+
  scale_x_discrete(labels = c("Wet-to-dry", "Mid-dry", "Dry-to-wet", "Mid-wet"))+
  theme_classic()+
  theme(legend.position = "none")

mod <- lmer(mean_mound_fluxCH4 ~ campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(mod)
Anova(mod)
report(mod)
emm <- emmeans(mod, ~ campaign)
pairs(emm)

# include species variation 
ggplot(all_fluxes_seasonal_avg, aes(x = campaign, y = mean_mound_fluxCH4, fill = species_s))+
  geom_boxplot()+
  ylab("CH4 flux umol/d/m2")+
  xlab("")+
  scale_x_discrete(labels = c("Wet-to-dry", "Mid-dry", "Dry-to-wet", "Mid-wet"))+ 
  scale_fill_manual(values = c("A. laurensis" = "#1b9e77", "C. acinaciformis" = "#d95f02", "N. magnus" =
                                 "#7570b3")) +
  theme_classic()

mod <- lmer(mean_mound_fluxCH4 ~ campaign + species_s + (1 | sample), data = all_fluxes_seasonal_avg)
summary(mod)
Anova(mod)
report(mod)
ggpredict(mod, terms = c("species_s", "campaign")) %>% plot(add.data = TRUE)

emm <- emmeans(mod, ~ campaign)
emm <- emmeans(mod, ~ species_s)
pairs(emm)

# b) test environmental predictors (temperature and precipitation) 
ggplot(all_fluxes_seasonal_avg, aes(x = temp, y = mean_mound_fluxCH4, color = campaign))+
  geom_point()+
  facet_wrap(~species_s)+
  geom_smooth(method = 'lm', se = TRUE)+
  ylab("CH4 flux umol/d/m2")+
  xlab("Average Mound Surface Temperature (C)")+
  scale_color_brewer(palette = "Dark2")+
  theme_classic()

mod <- lmer(mean_mound_fluxCH4 ~ temp*species_s*campaign + (1 | sample), data = all_fluxes_seasonal_avg)
Anova(mod)
summary(mod)

# model without all interactions
mod <- lmer(mean_mound_fluxCH4 ~ temp  + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(mod)
ggpredict(mod, terms = c("temp", "campaign", "species_s")) %>% plot(add.data = TRUE)

# nonlinear model 
residuals <- resid(mod)
fitted_values <- fitted(mod)
plot(fitted_values, residuals, main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")
# residuals dont look like theres a curve going on - try second order polynomial just to see what's going on 

```

non linear model exploration 
```{r}
# polynomial 
mod_poly <- lmer(mean_mound_fluxCH4 ~ poly(temp, 2) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(mod_poly)

residuals <- resid(mod_poly)
fitted_values <- fitted(mod_poly)

plot(fitted_values, residuals, main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")

qqnorm(resid(mod_poly))
qqline(resid(mod_poly))

# spl;ines
mod_spline <- lmer(mean_mound_fluxCH4 ~ ns(temp, 3) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(mod_spline)

residuals <- resid(mod_spline)
fitted_values <- fitted(mod_spline)

plot(fitted_values, residuals, main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")

qqnorm(resid(mod_spline))
qqline(resid(mod_spline))

AIC(mod, mod_poly, mod_spline)

# visualize non linear models 
# polynomial
new_data <- expand.grid(
  temp = seq(min(all_fluxes_seasonal_avg$temp), max(all_fluxes_seasonal_avg$temp), length.out = 100),
  species_s = unique(all_fluxes_seasonal_avg$species_s),
  campaign = unique(all_fluxes_seasonal_avg$campaign)
)
                       
new_data$predicted_flux <- predict(mod_poly, newdata = new_data, re.form = NA)

ggplot(new_data, aes(x = temp, y = predicted_flux, color = campaign)) +
  geom_line(aes(linetype = campaign)) +  # Different linetypes for campaigns
  facet_wrap(~ species_s) +              # Separate plots for each species
  labs(
    x = "Temperature (°C)", 
    y = "Predicted CH4 Flux (umol/d/m2)",
    title = "Polynomial Relationship between Temperature and CH4 Flux"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# splines
new_data$predicted_flux <- predict(mod_spline, newdata = new_data, re.form = NA)

ggplot(new_data, aes(x = temp, y = predicted_flux, color = campaign)) +
  geom_line(aes(linetype = campaign)) +  # Different linetypes for campaigns
  facet_wrap(~ species_s) +              # Separate plots for each species
  labs(
    x = "Temperature (°C)", 
    y = "Predicted CH4 Flux (umol/d/m2)",
    title = "Spline Relationship between Temperature and CH4 Flux"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


splines of different df 
```{r}
spline3 <- lmer(mean_mound_fluxCH4 ~ ns(temp, df = 3) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(spline3)
residuals <- resid(spline3)
fitted_values <- fitted(spline3)
plot(fitted_values, residuals, main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")
qqnorm(resid(spline3))
qqline(resid(spline3))

# how many df to use? seems like anything above 5 does not make much sense (lines start to get a bit weird)
spline4 <- lmer(mean_mound_fluxCH4 ~ ns(temp, df = 4) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(spline4)

spline5 <- lmer(mean_mound_fluxCH4 ~ ns(temp, df = 5) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(spline5)

spline6 <- lmer(mean_mound_fluxCH4 ~ ns(temp, df = 6) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(spline6)
residuals <- resid(spline6)
fitted_values <- fitted(spline6)
plot(fitted_values, residuals, main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")
qqnorm(resid(spline6))
qqline(resid(spline6))

spline7 <- lmer(mean_mound_fluxCH4 ~ ns(temp, df = 7) + species_s + campaign + (1 | sample), data = all_fluxes_seasonal_avg)
summary(spline7)

AIC(spline3, spline4, spline5, spline6, spline7, spline10)
```

Other interesting things 
--------------------------------
PROPORTIONAL FLUXES - from naked termite measurements, what proportion of the total flux is from termites vs mounds? 
```{r}
df <- p_fluxes_f[c("species", "prop_T_CH4", "prop_M_CH4", "prop_T_CO2", "prop_M_CO2")]
df_long <- df %>%
  pivot_longer(cols = starts_with("prop_"), 
               names_to = c(".value", "flux"), 
               names_pattern = "prop_(.*)_(.*)") %>%
  pivot_longer(cols = c(T, M), 
               names_to = "material", 
               values_to = "prop") 

ggplot(df_long[df_long$flux == "CH4",], aes(x = species, y = prop, fill = material))+
  geom_boxplot()+
  ylab("Proportion CH4 flux")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()

ggplot(df_long[df_long$flux == "CO2",], aes(x = species, y = prop, fill = material))+
  geom_boxplot()+
  ylab("Proportion CO2 flux")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()

# test what the ratio of termite:mound mass is - best if consistent across the df (seems to be largely true)
p_fluxes_f$termite_mound_ratio <- p_fluxes_f$mass_termite_g / p_fluxes_f$mass_mound_material_g
ggplot(p_fluxes_f, aes(x = species, y = termite_mound_ratio, fill = species))+
  geom_boxplot()+
  ylab("Termite mass:Mound mass")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()+
  theme(legend.position = "none")
```

Temperature variation by campaign
```{r}
# temperatures by campaign - this just shows seasonal variability
ggplot(all_fluxes_mound_avg, aes(x = campaign, y = temp, fill = campaign))+
  geom_boxplot()+
  geom_jitter()+
  ylab("Average Mound Surface Temperature (C)")+
  xlab("")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()+
  theme(legend.position = "none")

```

Code for daily cycle question
```{r}
# temperature changes throughout the day
t <- all_fluxes_mound[all_fluxes_mound$campaign %in% c("nov22", "aug23", "feb24"),]
t$flux_start <- strptime(t$flux_start, "%H:%M")

tt <- all_fluxes_mound[all_fluxes_mound$campaign %in% c("may22"),]
start <- nchar(tt$flux_start) - 7
tt$flux_start_test <- substring(tt$flux_start, start, nchar(tt$flux_start))
tt$flux_start <- strptime(tt$flux_start_test, "%H:%M")
tt <- tt[, -which(names(tt) == "flux_start_test")]

all_fluxes_withtime <- rbind(t, tt)

# temperature throughout the day by season
ggplot(all_fluxes_withtime, aes(x = as.POSIXct(flux_start), y = avg_respT, color = campaign))+
  geom_jitter()+
  geom_smooth(method = 'loess', se = TRUE)+
  ylab("Average Mound Surface Temperature (C)")+
  xlab("Time of day")+
  theme_classic()

# table of averages by season (all mounds, not those with repeat measurements)
detach(package:plyr)
season_avg <- all_fluxes_mound_avg %>% group_by(species_s, campaign) %>% 
  summarize(mean_CH4flux = mean(mean_mound_fluxCH4), 
            mean_CO2flux = mean(mean_mound_fluxCO2), 
            sd_CH4flux = sd(mean_mound_fluxCH4), 
            sd_CO2flux = sd(mean_mound_fluxCO2), 
            count = n())
```

Intramound flux variation 
```{r}
# INTRAMOUND VARIATION OF MOUND FLUX
all_fluxes_mound$position <- ifelse(all_fluxes_mound$position == "top ", "top", all_fluxes_mound$position)
ggplot(data = all_fluxes_mound, aes(x = species_s, y = flux.CH4, fill = position)) + 
  geom_boxplot()+ 
  theme_classic()+
  ylab("CH4 flux umol/d/m2")+
  ylim(0,24000)+ # removes big outlier
  xlab("")

# direction on mound
ggplot(data = all_fluxes_mound, aes(x = directon, y = flux.CH4, fill = directon)) + 
  geom_boxplot()+ 
  theme_classic()+
  ylab("CH4 flux umol/d/m2")+
  ylim(0,24000)+ # removes big outlier
  xlab("Direction")+
  scale_x_discrete(limits = c("E", "N", "S", "W"), labels = c("E", "N", "S", "W"))

# TIME OF DAY AND FLUX
# reorder levels of time of day
all_fluxes_daily$time_of_day <- factor(all_fluxes_daily$time_of_day, levels = c("morning", "midday", "afternoon", "night"))

# time by flux for species
ggplot(all_fluxes_daily, aes(x = species_s, y = flux.CH4, fill = time_of_day))+
  geom_boxplot()+
  theme_classic()+
  xlab("")+
  ylab("CH4 flux (umol/d/mound)")+
  ylim(0,25000) # removes outlier

# calculate mound average for flux and temp 
all_fluxes_daily$sampleTOD <- paste(all_fluxes_daily$sample, all_fluxes_daily$time_of_day, sep = "-")
daily_avg <- all_fluxes_daily %>% group_by(sampleTOD) %>% 
  summarise(mean_mound_fluxCH4=mean(flux.CH4),
            mean_mound_fluxCO2=mean(flux.CO2_umol), 
            mean_mound_fluxCH4_CO2=mean(CH4_CO2), 
            time_of_day = time_of_day, 
            species_s = species_s,
            sample = sample,
            temp = mean(avg_respT))
daily_avg_f <- distinct(daily_avg)

# temp by flux for species at mound-level average
ggplot(daily_avg_f, aes(x = temp, y = mean_mound_fluxCH4, color = sample))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_wrap(~species_s, scales = "free_y")+
  theme_classic()+
  xlab("Temperature (C)")+
  ylab("CH4 flux (umol/d/mound)")
```
