---
title: "2.prelim_fluxviz"
author: "abbey yatsko"
date: "26/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(viridis)
library(tidyr)
library(broom)
library(purrr)
library(tidyverse)
library(rstatix)
```

read in data 
```{r data load}
CO2_fluxfinal <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CO2_fluxfinal_all.csv")
CH4_fluxfinal<- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/CH4_fluxfinal_all.csv")
```

data clean and reorganize
```{r}
# merge in species data to flux data 
species <- read.csv("/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/Abbey termites .xlsx - Sheet1.csv")
species <- species[c("X", "ID_cleaned")]

# merge sp and CH4flux_final, CO2_fluxfinal
colnames(species)[1] = "sample"

CH4_fluxfinal <- merge(CH4_fluxfinal, species, by = c("sample")) 
CO2_fluxfinal <- merge(CO2_fluxfinal, species, by = c("sample")) 

# new column for methane:co2
# first need to convert CO2 from mmol to umol (in order to match CH4 in units of umol)
# (conversion factor 1000umol = 1mmol)
CO2_fluxfinal$flux.CO2_umol <- CO2_fluxfinal$flux.CO2*1000
CO2_fluxfinal <- CO2_fluxfinal %>% relocate(flux.CO2_umol, .after = flux.CO2)

# take out just CO2 flux to paste in working CH4_fluxfinal df
a <- CO2_fluxfinal[c("Time_start", "flux.CO2_umol")]
CH4_fluxfinal <- merge(CH4_fluxfinal, a, by = c("Time_start")) 
CH4_fluxfinal <- CH4_fluxfinal %>% relocate(flux.CO2_umol, .after = flux.CH4)

CH4_fluxfinal$ch4_co2 <-  CH4_fluxfinal$flux.CH4/CH4_fluxfinal$flux.CO2_umol
CH4_fluxfinal <- CH4_fluxfinal %>% relocate(ch4_co2, .after = flux.CH4)

# new column to add arboreal where 1 = arboreal, 0 = not arboreal
CO2_fluxfinal$arboreal[CO2_fluxfinal$chamber == "mini"] <- 1
CO2_fluxfinal$arboreal[!CO2_fluxfinal$chamber == "mini"] <- 0

CH4_fluxfinal$arboreal[CH4_fluxfinal$chamber == "mini"] <- 1
CH4_fluxfinal$arboreal[!CH4_fluxfinal$chamber == "mini"] <- 0

# subset df to create just methane from soils and just mounds (for later on. just to have df prepped)
soils <- CH4_fluxfinal[CH4_fluxfinal$flux_source == 's',]
mounds <- CH4_fluxfinal[CH4_fluxfinal$flux_source == 'm',]

# arboreal only 
arboreal <- mounds[mounds$arboreal == 1,]

# omit arboreal from mound measurements 
mounds <- mounds[mounds$arboreal == 0,]

# remove unidentified mounds 
mounds[mounds==""] <- NA
mounds_ID <- mounds[!is.na(mounds$ID_cleaned),]
mounds_ID$ID_cleaned <- as.factor(mounds_ID$ID_cleaned)

# remove duplicate ID columns 
# mounds_ID <- subset( mounds_ID, select = -c(ID_cleaned.x, ID_cleaned.y) )

# write.csv(mounds, "/Users/abbeyyatsko/Desktop/repos/TMI_flux/data/finalfluxes/mounds.csv")
```

methane flux for individual mounds, comparing seasons, by species
```{r flux}
# all mounds
a1 <- ggplot(data = mounds_ID, aes(x = sample, y = flux.CH4, fill = campaign)) + 
  geom_boxplot()+ 
  #facet_wrap(~ID_cleaned, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,18000)+ # removes outliers, makes it easier to see comparisons
  ylab("CH4 flux umol/d/m2")+
  labs(title = "mound flux")

# all soils
a2 <- ggplot(data = soils, aes(x = as.factor(sample), y = flux.CH4, fill = ID_cleaned)) +
  geom_boxplot()+
  geom_jitter()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylab("CH4 flux umol/d/m2")+
  labs(title = "soil")

# individual mounds with CH4:CO2
a3 <- ggplot(data = mounds_ID, aes(x = as.factor(sample), y = ch4_co2, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_wrap(~campaign, scales = "free_x")+
  ylim(0,0.05)+ # for outliers
  labs(title = "mound")

```

show significant mound methane flux to hardly any soil flux 
```{r by soil/mound}
b <- ggplot(data = CH4_fluxfinal, aes(x = as.factor(flux_source), y = flux.CH4)) + 
  geom_boxplot()+ 
  geom_jitter()+
  theme_light()
```

mound and soil methane flux by position
```{r by position}
# mounds 
c1 <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = flux.CH4, fill = position)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  xlab("Species")+
  scale_fill_brewer("Blues")+
  guides(fill=guide_legend(title="Position"))+
  theme_light()

c2 <- ggplot(data = mounds_ID, aes(x = as.factor(ID_cleaned), y = ch4_co2, fill = position)) + 
  geom_boxplot()+
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4:CO2")+
  xlab("Species")+
  scale_fill_brewer("Blues")+
  guides(fill=guide_legend(title="Position"))+
  ylim(0,0.07)+ # for outliers
  theme_light()

ggarrange(c1, c2, ncol = 1, common.legend = TRUE)

# soils (shows decreasing as you move from the mound, but all together insignificant flux)
c3 <- ggplot(data = soils, aes(x = as.factor(position), y = flux.CH4, fill = ID_cleaned)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  scale_fill_brewer("Blues")+
  theme_light()
```

methane flux for mounds based on cardinal direction of measurement
```{r by direction}
# all mounds, across different directions
d1 <- ggplot(data = mounds_ID, aes(x = as.factor(directon), y = flux.CH4)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  ylab("CH4 flux umol/d/m2")+
  theme_light()
d2 <- ggplot(data = mounds_ID, aes(x = as.factor(directon), y = ch4_co2)) + 
  geom_boxplot()+ 
  facet_wrap(~campaign, scales = "free_x")+
  theme_light()

ggarrange(d1, d2, ncol = 1)
```

methane flux by species throughout seasons 
```{r by species}
# compare species across campaigns 
all_data <- mounds_ID[!mounds_ID$ID_cleaned =="Macro",]
all_data$campaign <- as.factor(all_data$campaign)

ggplot(data = all_data, aes(x = campaign, y = flux.CH4, fill = as.factor(ID_cleaned))) + 
  geom_boxplot()+ 
  scale_fill_brewer()+
  ylab("CH4 flux (umol/d/m2)")+
  xlab("")+
  scale_x_discrete(limits = c("may22", "aug23", "nov22"))+
  theme_classic()

# compare campaigns for different species
ggplot(data = all_data, aes(x = as.factor(ID_cleaned), y = flux.CH4, fill = campaign)) + 
  geom_boxplot()+
  scale_fill_brewer()+
  ylab("CH4 flux (umol/d/m2)")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  theme_classic()

# just species effect 
ggplot(data = all_data, aes(x = as.factor(ID_cleaned), y = flux.CH4, fill = as.factor(ID_cleaned))) + 
  geom_boxplot()+ 
  scale_fill_brewer()+
  ylab("CH4 flux (umol/d/m2)")+
  xlab("")+
  scale_x_discrete(labels = c("A. laurensis", "C. acinaciformis", "N. magnus"))+
  theme_classic()+
  theme(legend.position = "none")

```

r2 distribution of fluxes 
```{r}
# co2 
mean(CO2_fluxfinal$R2.CO2)
ggplot(CO2_fluxfinal, aes(R2.CO2))+
  geom_histogram()+
  theme_classic()

# ch4 
mean(CH4_fluxfinal$R2.CH4)
ggplot(CH4_fluxfinal, aes(R2.CH4))+
  geom_histogram()+
  theme_classic()

weird <- CH4_fluxfinal[CH4_fluxfinal$R2.CH4 < 0.5,]
```


```{r}
# compare seasons between each species 
# first subset mounds by repeated seasonal measures (8 mounds / species)
nov <- mound_average[mound_average$campaign == "nov22",]
target_mounds <- unique(nov$sample)
season_contrast <- mound_average[mound_average$sample %in% target_mounds,]
season_contrast$species <- ifelse(season_contrast$ID_cleaned == "Ami", "A. laurensis", "")
season_contrast$species <- ifelse(season_contrast$ID_cleaned == "Nasuti", "N. magnus",season_contrast$species )
season_contrast$species <- ifelse(season_contrast$ID_cleaned == "Copto", "C. acinaciformis", season_contrast$species)

ggplot(season_contrast, aes(x = as.factor(species), y = mean_fluxCH4, fill = campaign)) + 
  geom_boxplot()+ 
  scale_fill_brewer("Blues", labels=c('Wet-to-dry', 'Dry-to-wet'))+
  ylab("CH4 flux umol/d/m2")+
  xlab("Species")+
  facet_wrap(~ID_cleaned,scales = "free_x" )+
  guides(fill=guide_legend(title=""))+
  stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 10000) +
  theme_classic()+
  theme(strip.text = element_blank())+
  ylab("CH4 flux (umol/d/m2)")+
  xlab("")
 # scale_f_manuel(labels= c("A. laurensis", "C. acinaciformis", "N.magnus"))

```

methane flux by mound average for different parts of the day 
```{r}
# merge in real time data from GPS records of mound being tagged (this is an overall mound estimate for the may campaign mounds)
maymounds <- mounds[mounds$campaign == "may22",]
maymounds_rm <- maymounds[maymounds$sample %in% target_mounds,]

novmounds <- mounds[mounds$campaign == "nov22",]
novmounds_rm <- novmounds[novmounds$sample %in% target_mounds,]

maytime <- read.csv("/Users/abbeyyatsko/Downloads/mound_gps (1).csv")
maytime <- maytime[c("name", "realtime")]
names(maytime)[1] <- 'sample'
maymounds <- merge(maymounds_rm, maytime, by = c("sample")) 
maymounds <- maymounds %>% relocate(realtime, .after = Time_start)
maymounds$realtime <- strptime(maymounds$realtime, "%H:%M")

novtime <- read.csv("/Users/abbeyyatsko/Downloads/novtimes.csv")
names(novtime)[2] <- 'realtime'
novmounds <- merge(novmounds_rm, novtime, by = c("sample")) 
novmounds <- novmounds %>% relocate(realtime, .after = Time_start)
novmounds$realtime <- strptime(novmounds$realtime, "%H:%M")

# plot for flux by time of day - may
ggplot(data = rbind(maymounds, novmounds), aes(x = avg_respT , y = flux.CH4)) + # as.POSIXct(realtime)
  geom_point()+
  geom_smooth(method = "lm")+ 
  #facet_wrap(~ID_cleaned)+
  #ylab("CH4 flux umol/d/m2")+
  guides(fill=guide_legend(title="Species"))+
  theme_classic() +
  theme(axis.text.x = element_text(size = 7)) +
  xlab("Average mound surface temperature (C)")+
  ylab("CH4 flux (umol/d/m2)")
  

summary(lm(avg_respT ~ as.POSIXct(realtime), data = rbind(maymounds, novmounds)[rbind(maymounds, novmounds)$ID_cleaned == "Copto",]))

summary(lm(flux.CH4 ~ avg_respT, data = rbind(maymounds, novmounds)))

# plot for flux by time of day - nov
ggplot(data = novmounds, aes(x = avg_respT, y = log(flux.CH4))) + 
  geom_point()+
  geom_smooth(method = "lm")+ 
  facet_wrap(~ID_cleaned, ncol =1)+
  #ylab("CH4 flux umol/d/m2")+
  guides(fill=guide_legend(title="Species"))+
  theme_classic() +
  theme(axis.text.x = element_text(size = 5)) 

summary(lm(flux.CH4 ~  avg_respT, data = novmounds[novmounds$ID_cleaned == "Ami",]))

```

methane flux by temperature 
```{r}
# scatter plot - methane 
h1 <- ggplot(data = mounds_ID[!mounds_ID$ID_cleaned == "Macro",], aes(x = avg_respT, y = flux.CH4, color = ID_cleaned)) + 
  geom_point(aes(color = ID_cleaned))+ 
  geom_smooth(method = 'loess', se = FALSE)+
  ylab("CH4 flux umol/d/m2")+
  xlab("Average Mound Surface Temperature (C)")+
  scale_color_brewer("Blues") +
  #facet_wrap(~campaign)+ # take this out to look at overall temperature effect
  guides(color=guide_legend(title="Campaign"))+
  theme_linedraw()

# lm by species for both may/nov
may <- mounds_ID[mounds_ID$campaign == "may22",]
nov <- mounds_ID[mounds_ID$campaign == "nov22",]

summary(lm(flux.CH4 ~ avg_respT, data = mounds_ID))

# run linear model
mounds_ID %>% 
  group_by(ID_cleaned) %>% 
  group_split() %>% 
  map_dfr(.f = function(mounds_ID){
    lm(flux.CH4 ~ avg_respT, data = mounds_ID) %>% 
      glance() %>% 
      add_column(SPECIES = unique(mounds_ID$ID_cleaned), .before=1)
  })

h2 <- ggplot(data = mounds_ID, aes(x = avg_respT, y = flux.CO2_umol)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  ylab("CO2 flux umol/d/m2")+
  xlab("Average Mound Surface Temperature (C)")+
  facet_wrap(~ID_cleaned, scales = "free_x")+
  scale_color_brewer("Blues") +
  guides(color=guide_legend(title="Campaign"))+
  theme_linedraw()

# run linear model
mounds_ID %>% 
  group_by(ID_cleaned) %>% 
  group_split() %>% 
  map_dfr(.f = function(mounds_ID){
    lm(flux.CO2_umol ~ avg_respT, data = mounds_ID) %>% 
      glance() %>% 
      add_column(SPECIES = unique(mounds_ID$ID_cleaned), .before=1)
  })

h3 <- ggplot(data = mounds_ID, aes(x = avg_respT, y = ch4_co2)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  ylab("CH4:CO2")+
  xlab("Average Mound Surface Temperature (C)")+
  facet_wrap(~ID_cleaned, scales = "free_x")+
  scale_color_brewer("Blues") +
  guides(color=guide_legend(title="Campaign"))+
  theme_linedraw()

# run linear model
mounds_ID %>% 
  group_by(ID_cleaned) %>% 
  group_split() %>% 
  map_dfr(.f = function(mounds_ID){
    lm(ch4_co2 ~ avg_respT, data = mounds_ID) %>% 
      glance() %>% 
      add_column(SPECIES = unique(mounds_ID$ID_cleaned), .before=1)
  })

ggarrange(h1, h2, h3, ncol = 1, common.legend = TRUE)

# average temperature by mound via campaign
mounds_ID$mound_campaign <- paste(mounds_ID$sample,mounds_ID$campaign, sep = "-")
x <- aggregate(mounds_ID$avg_respT, list(mounds_ID$mound_campaign), FUN=mean)
colnames(x)[1] = "mound_campaign"
colnames(x)[2] = "mound_temp"
mounds_ID <- merge(mounds_ID, x, by = c("mound_campaign")) 
str(mounds_ID)
mounds_ID$campaign <- as.factor(mounds_ID$campaign)

h4 <- ggplot(data = mounds_ID, aes(x = mound_temp, y = flux.CH4, fill = as.factor(sample))) + 
  geom_boxplot()+ 
  ylab("CH4 flux umol/d/m2")+
  facet_grid(campaign~ID_cleaned)+
  theme(legend.position="none")+
  scale_fill_viridis(discrete = TRUE, option = "D")+
  ylim(0,16000) # for outliers

```

scaled mound flux from calculated SA
```{r}
# find average flux by mound, calculating seperately remeasured mounds  
# group by mound_campaign
y1 <- mounds_ID %>% group_by(mound_campaign) %>% 
  summarise(mean_fluxCH4=mean(flux.CH4),mean_fluxCO2=mean(flux.CO2_umol), mean_fluxCH4CO2=mean(ch4_co2) )
y1 <- separate(y1, col=mound_campaign, into=c('sample', 'campaign'), sep='-')

# merge with mound_SA
mound_SA <- read.csv("/Users/abbeyyatsko/Downloads/mound SA - Sheet1 (1).csv")
mound_SA <- mound_SA[, c('sample', 'working_sa_m2')]
mound_SA <- mound_SA[!is.na(mound_SA$working_sa_m2),]
z <- merge(y1, mound_SA, by = "sample" )
z$totalCH4flux <- z$mean_fluxCH4 * z$working_sa_m2
z$totalCO2flux <- z$mean_fluxCO2 * z$working_sa_m2
z <- merge(z, species, by = "sample" )

# graph mound flux by mound size (use scaled flux - this is the actual flux being represented by the true size of the mound)
# methane
z$species <- ifelse(z$ID_cleaned == "Ami", "A. laurensis", "")
z$species <- ifelse(z$ID_cleaned == "Nasuti", "N. magnus",z$species )
z$species <- ifelse(z$ID_cleaned == "Copto", "C. acinaciformis", z$species)
ggplot(data = z[!z$ID_cleaned == "Macro",], aes(x = working_sa_m2, y = log(totalCH4flux))) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("adj.R2", "f", "p"))) +
  geom_point()+ 
  geom_smooth(method = 'lm')+
  scale_color_brewer("Blues")+
  facet_wrap(~species, ncol = 1)+ # scales = "free_x"
  theme_classic()+
  ylab("Total mound CH4 flux (log scale, umol/d/mound)") +
  xlab("Mound surface area (m2)")

summary(lm(log(totalCH4flux) ~  working_sa_m2, data = z[z$ID_cleaned == "Nasuti",]))

# carbon dioxide
ggplot(data =  z[z$ID_cleaned == "Ami",], aes(x = working_sa_m2, y = totalCO2flux, color = campaign)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  facet_wrap(~ID_cleaned, scales = "free_x")+ 
  theme_classic()+
  ylab("Total mound CO2 flux umol/d/mound") +
  xlab("mound surface area")

# methane:carbon dioxide
ggplot(data = z[z$ID_cleaned == "Nasuti",], aes(x = working_sa_m2, y = mean_fluxCH4CO2, color = campaign)) + 
  geom_point()+ 
  geom_smooth(method = 'lm')+
  facet_wrap(~ID_cleaned, scales = "free_x")+
  theme_classic()+
  ylab("Total mound CH4:CO2") +
  xlab("mound surface area")
```

calculate average flux per species 
```{r}
# average fluxes - may 
# remove NAs 
maymounds_ID <- maymounds[!is.na(maymounds$ID_cleaned),]

may_avg <- maymounds_ID %>% group_by(ID_cleaned) %>% 
  summarise(mean_flux=mean(flux.CH4),
            .groups = 'drop')

# average fluxes - nov 
nov_avg <- novmounds %>% group_by(ID_cleaned) %>% 
  summarise(mean_flux=mean(flux.CH4),
            .groups = 'drop')

# to convert umol/m2*d to ug/m2*d:
# avgflux_ami <- avgflux_ami*(1/1000000)*(16.04)*(1000000) # molecular mass conversion to g, then conversion back to ug for comparison 
# avgflux_nasute <- avgflux_nasute*(1/1000000)*(16.04)*(1000000) 
# avgflux_copto <- avgflux_copto*(1/1000000)*(16.04)*(1000000) 
# final units for average fluxes are now in ug CH4/m2*d
```

